Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-08-24 manuscript.Rnw
Minor edits throughout

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation \citep{ise2010, dekauwe2014,friend2014,ise2010}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \citep{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass \citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014,negron2015}.

@@

In this paper we are primarily interested in the distribution of biomass ('partitioning') between leaves and woody stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and individual plants \citep{poorter2012,poorter2015}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Several classic, oft-cited, studies have assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.

Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because smaller-scale comparisons have shown responsiveness of biomass partitioning to environmental drivers \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. A recent study by \citet{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to.

Despite considerable advances in theory underlying allometric scaling in plants \citep{enquist2003, west1999, savage2010} we do not yet have a clear understanding of potential differences between PFTs in terms of biomass partitioning. Instead, most previous work has focussed on understanding size-related shifts in biomass partitioning, as governed by constraints including hydraulic supply and mechanical stability \citep{savage2010}. Rather than to advance a new model of biomass partitioning or allocation, we take the view that a broader evidence base is needed to elucidate patterns between PFTs. This should enable those building GVMs to refine algorithms and parameter values to more closely match the observations. Here, we use a unique, new database \citep{falster2015} (Fig.~\ref{fig:figoverview}) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. A recent compilation of plant biomass data \citep{poorter2015} speculated that differences between gymnosperms and angiosperms in distribution of biomass between leaves and stems is related to differences in leaf mass per area. Here we can directly test this hypothesis as our database, unlike that of \citet{poorter2015}, includes many observations of leaf area as well as leaf mass measured on the same plants.

A second objective was to study relationships between biomass partitioning and climate variables or biome at a global scale. The dataset includes observations of biomass and size metrics for individual plants, compiled from 175 studies across nine vegetation types (Fig. S1), across the three major biomes (boreal, temperate and tropical). In this paper we focus on field-grown woody plants, spanning the entire size range of woody plants (0.01 - \textgreater 100 m height). One key challenge is thus to account for the very large size variation commonly found in any allometric variable \citep{niklas1994}. We do this by fitting a statistical model that does no assume a functional dependence on plant size, allowing us to study PFT patterns at a common plant height.

@@

where $M_i$ is the biomass in some compartment (leaves, stems or roots), $a_i$ the annual allocation of NPP to that compartment, and $\tau_i$ the annual turnover of compartment $i$. It is thus easy to see that partitioning ($M_i/M_T$, where $M_T$ is total biomass) can be different from allocation ($a_i$) because turnover ($\tau_i$) differs between leaf and wood biomass. Here, we present data on biomass partitioning, which can inform models of allocation only when estimates of turnover are available.

@@

We used the recently compiled Biomass And Allometry Database (BAAD) \citep{falster2015}, which in total include records for \Sexpr{nrow(baad_all$data)} individuals. The database does not overlap with the recent large compilation of \citet{poorter2015}. In this paper we restrict our analysis to records that include leaf mass ($M_F$), leaf area ($A_F$), above-ground woody biomass ($M_S$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (typically 1.3m) ($A_{Sbh}$) (n=\Sexpr{nrow(dataset)}). The database contains many more variables, for example root biomass for a much smaller subset of studies. Here we limit the analysis to patterns in aboveground biomass distribution. For each analysis, we used different subsets of the data because not all variables were measured in each study. Sample sizes by PFT are summarized in Table 1. We excluded glasshouse and common garden studies, including only field-grown woody plants (including natural vegetation, unmanaged and managed plantations). We considered three PFTs : evergreen angiosperms, evergreen gymnosperms, and deciduous angiosperms. We excluded  deciduous gymnosperms because few data were available. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes. To assess the coverage of the global climate space by the dataset, we extracted mean annual temperature and precipitation from Worldclim \citep{worldclim}, excluding areas without woody vegetation (taken from the global land cover database GLC-SHARE \citep{lathamglobal}).

@@

$$a = c_0H^{c_1.}$$

@@

We used generalised additive models (GAM) to visualize the relationships between biomass and plant size variables, and to estimate variables and their confidence intervals such as LMF at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalized regression smoother \citep{wood2006} to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect. We used the \emph{mgcv} package in R \citep{wood2006}. All variables (except MAP and MAT) were log-transformed prior to analysis.

@@

Despite the size of our dataset, the strong size-dependence of LMF greatly hinders comparisons across climatic gradients, due to the small sample size within each species or site. To study climate effects on biomass partitioning, we therefore further decomposed LMF as the product,

@@

This decomposition showed that the vast majority of size-related variation was captured by $A_S/M_T$ alone. Similarly, we decomposed LAR as the product of $A_F/A_S$ and $A_S/M_T$. We analyzed climate effects on $M_F/A_S$ and $A_F/A_S$ in two ways: with quantitative variables (mean annual precipitation, MAP; mean annual temperature, MAT), and by biome (tropical / temperate/ boreal), a simple classification taking into account both MAT and MAP (Fig. S1). In both cases we used PFT and plant height as covariates.

@@

The raw data in Fig.~\ref{fig:figrawdata} show a steeper increase of woody aboveground biomass ($M_S$) with plant height compared to foliage biomass ($M_F$). As a result, LMF decreased with plant height, with the three plant functional types clearly differing in LMF across nearly the entire size range (Fig.~\ref{fig:figlmf}a). When corrected for plant height by estimating LMF at a common plant height, we found that LMF was proportional to the average leaf mass per area (LMA) across the three PFTs (Fig.~\ref{fig:figlmf}b). As a consequence, LAR was invariant between PFTs, because LAR = LMF / LMA (Fig. ~\ref{fig:figlmf}c and Fig. S2).

Mirroring the results for LMF, we found that the amount of leaf mass per unit stem area ($M_F/A_S$) differs among the three PFTs, that these differences are also proportional to LMA (Fig. ~\ref{fig:figlmf}d), and that PFTs are similarly invariant in the amount of leaf area per unit stem area ($A_F/A_S$) (Fig.~\ref{fig:figlmf}e).

The above results highlight how LMA drives differences in biomass partitioning among dominant woody PFTs. To assess whether LMA could replace PFT when predicting biomass partitioning, we fit a linear mixed-effects model to LMF and $M_F/A_S$ using PFT and plant height (and the quadratic term) as predictors (and all interactions). We then replaced PFT with LMA, and found that the model with LMA could explain almost as much variation in LMF as the model with PFT ($R^2$ = 0.74 with PFT vs. 0.62 with LMA) and $M_F/A_S$ ($R^2$ = 0.29 vs. 0.28).

There was considerable variation in all studied variables between species within PFTs (Fig.\ref{fig:fighist}). To understand whether we can improve on a PFT-based classification by including other traits that affect biomass partitioning, we decomposed $M_F/A_S$ into LMA ($M_F/A_F$), and the ratio of leaf area to stem cross-sectional area ($A_F/A_S$) (Eq. \ref{eq:M2}). The ratio $M_F/A_S$ is relevant because plant leaf mass is frequently estimated from $A_S$ with allometric equations \citep{shinozaki1964b, chave2005}, using records of stem diameter commonly recorded on long-term monitoring plots.

@@

The correlation between $M_F/A_S$ and LMA was found to hold also within PFTs (Fig.~\ref{fig:figlmfspecies}a), explaining 30\% of the variation in $M_F/A_S$ across species. The regressions across species by PFT were broad and overlapping, with a more general relationship extending across the entire LMA axis. A larger fraction of the variation in $M_F/A_S$ was explained by species-level differences in $A_F/A_S$ (Fig.~\ref{fig:figlmfspecies}b). As previously noted, $A_F/A_S$ does not differ systematically among PFTs (Fig.~\ref{fig:figlmf}e), but does vary close to two orders of magnitude across species. In Fig.~\ref{fig:figlmfspecies}b, the separation among PFTs in $M_F/A_S$ arises due to differences in LMA.

We found that decomposing LMF and LAR as shown in Eq. \ref{eq:M1} was very useful to study climate effects because the second term ($A_S/M_T$) absorbs nearly all of the size related variation in LMF (Table 2 and Fig. S3) and is otherwise fairly conserved across PFTs (Fig.~\ref{fig:fighist}). As shown in Fig.~\ref{fig:figlmfspecies}b and Fig.~\ref{fig:figlmf}d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Table 2). $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analyzed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig.~\ref{fig:figclimate}). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table 2). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig.~\ref{fig:figclimate}b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table 2).

@@

We found that, as expected, plant size strongly influences LMF (Fig.~\ref{fig:figlmf}) and LAR (Fig. S2) (see also \citet{poorter2012,poorter2015}). It is thus necessary to correct for plant size when comparing biomass partitioning parameters, as has been noted many times \citep{mcconnaughay1999}. We used a semi-parametric approach to account for plant size, which has the advantage that it does not require an \emph{a priori} assumption on the functional relationship. This was useful because both LMF and LAR showed very non-linear patterns with plant height, even on a logarithmic scale, consistent with \citet{poorter2015}.

@@

We found that patterns in $M_F/A_S$ across PFTs mirrored those of LMF (Fig.~\ref{fig:figlmf}). Because LMF may be decomposed into the terms $M_F/A_S$ and $M_T/A_S$ (\ref{eq:M2}), this result suggests that $M_T/A_S$ does not vary between PFTs (see also Fig. S3). If this result holds, it means that aboveground biomass may be estimated across PFTs with a single equation that does not differ between PFTs. For tropical forests, \citet{chave2005} showed that aboveground biomass can be well estimated as,

@@

where $F$ is a form factor, and $\rho$ wood density (kg m$^3$). After rearranging, we find that this equation predicts that $M_T/A_S$ is proportional to plant height. We found some support for this prediction (Fig. S3). It is timely to use database made public by \citet{falster2015} (BAAD) and \citet{poorter2015} to develop more general equations to estimate individual plant aboveground biomass across the globe.

We found that climate did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for this difference is that our measurements were taken on individual plants, whereas in  \citet{reich2014} they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig.~\ref{fig:fighist}), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Table 2).

@@

Our sincere thanks to everyone who contributed data to the Biomass and Allometry Database. We thank Martin de Kauwe for discussions and comments on an earlier version of this paper. Figure 2a uses an image by Lana Heydon, IAN Image Library http://ian.umces.edu/imagelibrary.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS2.pdf}
    \caption{Raw data for leaf biomass ($M_F$) and total above-ground woody biomass ($M_S$) for each of the PFTs, as a function of total plant height ($H$). Each point is an individual plant. Sample sizes are listed in Table 1}
    \label{fig:figrawdata}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS5.pdf}
    \caption{Plant functional types diverge strongly in mass-based partitioning, but converge in area-based partitioning. Shown are histograms (as probability density functions) of leaf mass per area ($M_F/A_F$), leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/M_S$) grouped by the three PFTs. Arrows indicate means by PFT.}
    \label{fig:fighist}
\end{figure}

>>>>

The partitioning of forest biomass among leaves and stems strongly influences the productivity and carbon cycle of the world's vegetation \citep{ise2010, dekauwe2014,friend2014}. Biomass stored in woody stems has a long residence time \citep{luyssaert2008}, whereas leaf biomass turns over quickly, entering the soil carbon cycle where the majority of carbon is released back to the atmosphere \citep{ryanlaw2005}. Globally, forests store approximately 360Pg of carbon in living biomass \citep{panetal2011}, equivalent to almost 40 years of current anthropogenic CO\textsubscript{2} emissions \citep{friedlingsteinetal2014}. Reducing uncertainties about biomass partitioning and carbon residence times in GVMs is therefore a priority for understanding the effects of climate and other environmental change on the global carbon cycle \citep{friend2014,dekauwe2014,negron2015}.

@@

In this paper we are primarily interested in the distribution of biomass ('partitioning') between leaves and woody stems, an important component of the residence time of carbon in ecosystems \citep{friend2014}. Previous work based on either whole stands \citep{oneilldeangelis1981,enquistniklas2002,reich2014} and or a mix on stand-- and individual-based measurements \citep{poorter2012, poorter2015}, reported differences between angiosperms and gymnosperms in the amount of leaf biomass per unit above-ground biomass (the 'leaf mass fraction', LMF). \citet{poorter2012} and \citet{enquistniklas2002} attributed higher LMF in gymnosperms to longer leaf lifespan (LL) compared to typical angiosperms, but this begs the question whether differences in LMF are also apparent between deciduous and evergreen functional types within angiosperms. It is also unknown whether PFTs with higher leaf biomass also have higher total leaf area, which is relevant because leaf area drives total light interception and thus productivity. Some oft-cited studies have also assumed that gymnosperms carry more leaf area than angiosperms \citep{chabothicks1982, bond1989}.

Little is known about global-scale patterns in LMF and LAR in relation to climate. It can be expected that biomass partitioning is correlated with precipitation or mean annual temperature because smaller-scale comparisons have shown responsiveness of biomass partitioning to environmental drivers \citep{berninger1994,callaway1994,delucia2000,poyatos2007}. Moreover, a recent study by \citet{reich2014} demonstrated that stand-scale biomass partitioning was related to mean annual temperature across diverse forest stands. Indeed, many GVMs assume that climate affects biomass partitioning within a given PFT, often through soil water stress or other abiotic stress factors (see Notes S1). Again, however, these models do not agree on the degree of plasticity in biomass partitioning, or which climate variables it should respond to.

Despite considerable advances in theory underlying allometric scaling in plants \citep{enquist2003, west1999, savage2010} we do not yet have a clear understanding of potential differences between PFTs in terms of biomass partitioning. Instead, most previous work has focussed on understanding size-related shifts in biomass partitioning, as governed by constraints including hydraulic supply and mechanical stability \citep{savage2010}. Rather than to advance a specific model of biomass partitioning or allocation, our view is that a broader evidence base is needed first to elucidate patterns between PFTs. This should enable those building GVMs to refine algorithms and parameter values to more closely match the observations. Here, we use a unique, new database \citep{falster2015} (Fig.~\ref{fig:figoverview}) to establish general rules on how biomass partitioning differs among three dominant woody PFTs across the globe: evergreen gymnosperms, evergreen angiosperms, and deciduous angiosperms. A recent compilation of plant biomass data \citep{poorter2015} speculated that differences between gymnosperms and angiosperms in distribution of biomass between leaves and stems is related to differences in leaf mass per area. Here we can directly test this hypothesis as our database, unlike those of \citet{enquistniklas2002}, \citet{reich2014}, and \cite{poorter2015}, includes many observations of leaf area as well as leaf mass measured on the same plants.

A second objective was to study relationships between biomass partitioning and climate variables or biome at a global scale. The dataset includes observations of biomass and size metrics for individual plants, compiled from 175 studies across nine vegetation types (Fig. S1), across the three major biomes (boreal, temperate and tropical). In this paper we focus on field-grown woody plants, spanning the entire size range of woody plants (0.01 - \textgreater 100 m height). One key challenge is thus to account for the very large size variation commonly found in any allometric variable \citep{niklas1994}. We do this by fitting a semi-parametric data-driven statistical model, i.e. one that does not assume a particular functional form, thereby allowing us to study PFT patterns at a common plant height.

@@

where $M_i$ is the biomass in some compartment (leaves, stems or roots) remaining on the plant, $a_i$ the annual allocation of NPP to that compartment, and $\tau_i$ the annual turnover (or loss) of compartment $i$ from the plant. It is thus easy to see that partitioning ($M_i/M_T$, where $M_T$ is total biomass) can be different from allocation ($a_i$) because turnover ($\tau_i$) differs between leaf and wood biomass. Here, we present data on biomass partitioning, which can inform models of allocation only when estimates of turnover are available.

@@

We used the recently compiled Biomass And Allometry Database (BAAD) \citep{falster2015}, which in total includes records for \Sexpr{nrow(baad_all$data)} individuals. The database has limited overlap with the recent large compilation of \citet{poorter2015} and differs in that a majority of measurements for large plants are individual rather than stand-based. In this paper we restrict our analysis to records that include leaf mass ($M_F$), leaf area ($A_F$), above-ground woody biomass ($M_S$), plant height ($H$), and stem area measured at ground level ($A_S$), or at breast height (typically 1.3m) ($A_{Sbh}$) (n=\Sexpr{nrow(dataset)}). The database contains many more variables, for example root biomass for a much smaller subset of studies. Here we limit the analysis to patterns in aboveground biomass distribution. For each analysis, we used different subsets of the data because not all variables were measured in each study. Sample sizes by PFT are summarised in Table \ref{table:counttable}. We excluded glasshouse and common garden studies, including only field-grown woody plants (including natural vegetation, unmanaged and managed plantations). We considered three PFTs : evergreen angiosperms, evergreen gymnosperms, and deciduous angiosperms. We excluded  deciduous gymnosperms because few data were available. All locations were further separated into boreal (including sub-boreal), temperate, and tropical biomes. To assess the coverage of the global climate space by the dataset, we extracted mean annual temperature and precipitation from Worldclim \citep{worldclim}, excluding areas without woody vegetation (taken from the global land cover database GLC-SHARE \citep{lathamglobal}).

@@

$$a = c_0H^{c_1}.$$

@@

We used generalised additive models (GAM) to capture  the relationships between biomass and plant size variables, and to estimate variables and their confidence intervals such as LMF at a common plant height. In all fitted GAMs, we used a cubic regression spline. For the smoothed term in the model (plant height), we used up to 3 or 4 degrees of freedom, which resulted in biologically realistic smoothed relationships. Within the GAM, we used a penalised regression smoother \citep{wood2006} to allow the final degree of smoothness to be estimated from the data. In all fitted GAMs, we used species-dataset combination as a random effect. All variables (except MAP and MAT) were log-transformed prior to analysis.

@@

Despite the exceptional size of our dataset, the strong size-dependence of LMF still hinders comparisons across climatic gradients, due to the small sample size available within each species or site when sampling at a common height. To study climate effects on biomass partitioning, we therefore further decomposed LMF as the product,

@@

This decomposition showed that the vast majority of size-related variation was captured by $A_S/M_T$ alone, indicating that comparisons among $M_F/A_S$ could be made without needing to compare at a common height. A similar decomposition  of LAR as the product of $A_F/A_S$ and $A_S/M_T$ produced the same outcome. We therefore analysed for climatic effects on $M_F/A_S$ and $A_F/A_S$ in two ways: with quantitative variables (mean annual precipitation, MAP; mean annual temperature, MAT), and by biome (tropical / temperate/ boreal), a simple classification taking into account both MAT and MAP (Fig. S1). In both cases we used PFT and plant height as covariates varying across climate space.

All analyses were conducted in R v3.2.0 \citep{R-2015}. GAMs were fitted using \emph{mgcv} package \citep{wood2006}. The code to replicate this analysis are available on GitHub at http://github.com/RemkoDuursma/baadanalysis.

@@

The raw data in Fig.~\ref{fig:figrawdata} show a steeper increase of woody aboveground biomass ($M_S$) with plant height compared to foliage biomass ($M_F$). As a result, LMF decreased with plant height, with the three plant functional types clearly differing in LMF across nearly the entire size range (Fig.~\ref{fig:figlmf}a). When we corrected for plant height by estimating LMF at a common plant height, we found that LMF was proportional to the average leaf mass per area (LMA) across the three PFTs (Fig.~\ref{fig:figlmf}b). As a consequence, LAR was invariant between PFTs, because LAR = LMF / LMA (Fig. ~\ref{fig:figlmf}c and Fig. S2).

Mirroring the results for LMF, we found that the amount of leaf mass per unit stem area ($M_F/A_S$) differed among the three PFTs, that these differences were also proportional to LMA (Fig. ~\ref{fig:figlmf}d), and that PFTs were similarly invariant in the amount of leaf area per unit stem area ($A_F/A_S$) (Fig.~\ref{fig:figlmf}e).

The above results highlight how LMA drives differences in biomass partitioning among dominant woody PFTs. To assess whether LMA could replace PFT when predicting biomass partitioning, we fit a linear mixed-effects model to LMF and $M_F/A_S$ using PFT and plant height (and the quadratic term) as predictors (and all interactions). We then replaced PFT with LMA, and found that the model with LMA could explain almost as much variation in LMF as the model with PFT ($R^2$ = 0.74 with PFT vs. 0.62 with LMA), likewise for $M_F/A_S$ ($R^2$ = 0.29 vs. 0.28).

There was considerable variation in all studied variables between species within PFTs (Fig. ~\ref{fig:fighist}). To understand whether we can improve on a PFT-based classification by including other traits that affect biomass partitioning, we decomposed $M_F/A_S$ into LMA ($M_F/A_F$), and the ratio of leaf area to stem cross-sectional area ($A_F/A_S$):

@@

The ratio $M_F/A_S$ is relevant because plant leaf mass is frequently estimated from $A_S$ with allometric equations \citep{shinozaki1964b, chave2005}, using records of stem diameter commonly recorded on long-term monitoring plots.  The correlation between $M_F/A_S$ and LMA was found to hold also within PFTs (Fig.~\ref{fig:figlmfspecies}a), explaining 30\% of the variation in $M_F/A_S$ across species. The regressions across species by PFT were broad and overlapping, with a more general relationship extending across the entire LMA axis. A larger fraction of the variation in $M_F/A_S$ was explained by species-level differences in $A_F/A_S$ (Fig.~\ref{fig:figlmfspecies}b). As previously noted, $A_F/A_S$ does not differ systematically among PFTs (Fig.~\ref{fig:figlmf}e), but does vary close to two orders of magnitude across species. In Fig.~\ref{fig:figlmfspecies}b, the separation among PFTs in $M_F/A_S$ arises due to differences in LMA.

We found that decomposing LMF and LAR as shown in Eq. \ref{eq:M1} was very useful to study climate effects because the second term ($A_S/M_T$) absorbed nearly all of the size-related variation in LMF (Table \ref{table:statstable} and Fig. S3) and is otherwise fairly conserved across PFTs (Fig. S3). As shown in Fig.~\ref{fig:figlmfspecies}b and Fig.~\ref{fig:figlmf}d, the term $M_F/A_S$ exhibits comparable differences between PFTs to LMF, but unlike LMF, $M_F/A_S$ is nearly independent of plant height (Table \ref{table:statstable}). $M_F/A_S$ is thus a useful proxy for LMF that can be compared across species and sites.

Regardless of how we analysed the data, we found only weak climatic effects on biomass partitioning within PFTs (Fig.~\ref{fig:figclimate}). Biome consistently explained very little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by only 0.01 - 0.06, see Table \ref{table:statstable}). Likewise, we found weak and inconsistent effects of climate variables (MAP and MAT) (Fig.~\ref{fig:figclimate}b-c). These variables explained little variation when added to a statistical model in addition to PFT and plant height ($R^2$ increased by 0.01 and 0.11, respectively, see Table \ref{table:statstable}).

@@

We found that, as expected, plant size strongly influences LMF (Fig.~\ref{fig:figlmf}) and LAR (Fig. S2) (see also \citet{poorter2012,poorter2015}). It is thus necessary to correct for plant size when comparing biomass partitioning parameters, as has been noted many times \citep{mcconnaughay1999}. We used a semi-parametric approach to account for plant size, which has the advantage that it does not require an \emph{a priori} assumption on the functional relationship. This was useful because both LMF and LAR showed very non-linear patterns with plant height, even on a logarithmic scale, consistent with recent other results \citep{poorter2015}.

@@

We found that patterns in $M_F/A_S$ across PFTs mirrored those of LMF (Fig.~\ref{fig:figlmf}). Because LMF may be decomposed into the terms $M_F/A_S$ and $M_T/A_S$ (eq. \ref{eq:M2}), this result suggests that $M_T/A_S$ does not vary between PFTs (see also Fig. S3). If this result holds, it means that aboveground biomass may be estimated across PFTs with a single equation that does not differ between PFTs. For tropical forests, \citet{chave2005} showed that aboveground biomass can be estimated as,

@@

where $F$ is a form factor, and $\rho$ wood density (kg m$^3$). After rearranging, this equation predicts that $M_T/A_S$ is proportional to plant height. We found some support for this prediction (Fig. S3), suggesting further improvements could be made by incorporating this dependence. It is timely to use database made public by \citet{falster2015} (BAAD) and \citet{poorter2015} to develop more general equations to estimate individual plant aboveground biomass across the globe.

We found that climate did not appreciably affect biomass partitioning between leaves and stems. This finding is in contrast with a recent study where LMF was found to increase with MAT across the globe \citep{reich2014}. One explanation for this difference is that our measurements were taken on individual plants, whereas in  \citet{reich2014} they were on whole stands. This suggests climatic effects on stand-level biomass partitioning occur primarily by altering the size-distribution and stand density (number of individuals per unit area), rather than partitioning within individual plants. With the caveats that there is still variation between species and studies unaccounted for (Fig.~\ref{fig:fighist}), and that coarse climate variables such as MAT and MAP may mask small-scale climate effects on biomass partitioning, our results strongly suggest the effects of PFT are stronger than any direct climate effect on biomass partitioning within individual plants (Table \ref{table:statstable}).

@@

Our sincere thanks to everyone who contributed data to the Biomass and Allometry Database. We thank Martin De Kauwe and anonymous referees for comments on an earlier version of this paper. Thanks also to Rich FitzJohn for advice about using the remake package. Figure 2a uses an image by Lana Heydon, IAN Image Library http://ian.umces.edu/imagelibrary.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS2.pdf}
    \caption{Raw data for leaf biomass ($M_F$) and total above-ground woody biomass ($M_S$) for each of the PFTs, as a function of total plant height ($H$). Each point is an individual plant. Sample sizes are listed in Table \ref{table:counttable}}
    \label{fig:figrawdata}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureS5.pdf}
    \caption{Plant functional types diverge strongly in mass-based partitioning, but converge in area-based partitioning. Shown are histograms (as probability density functions) of leaf mass per area ($M_F/A_F$), leaf mass per unit basal stem area ($M_F/A_S$) and leaf area per unit basal stem area ($A_F/M_S$) grouped by the three PFTs. Arrows indicate means by PFT. Colours as in Fig. \ref{fig:figlmf}.}
    \label{fig:fighist}
\end{figure}
